Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2020-01-26 write_up/mixfishsim_ecolmod_revisions.tex
Incorparting Coilin and Jean-Marc's comments

\usepackage[draft]{changes}%\usepackage[draft]{changes}

@@

%%
To understand how data resolution impacts inference on mixed fisheries
interactions we developed a highly resolved spatiotemporal discrete-event
simulation model (`MixFishSim')\textcolor{blue}{[CM: I think \emph{MixFishSim} is more effective]} incorporating: i) delay-difference population
dynamics, ii) population movement using Gaussian Random Fields to simulate
patchy, heterogeneously distributed \added{and moving fish} populations, and
iii) fishery dynamics for multiple fleet characteristics based on population
targeting under an explore-exploit strategy.\deleted{ This is implemented via a mix of
correlated random walk movement (for exploration) and learned behaviour (for
exploitation) phases of the fisheries. }\\ 
%%
\deleted{Fifty years of sub-daily fishing activity was simulated and used to
	draw inference on the underlying community structures. W}\added{We
	applied `MixFishSim' to\deleted{draw}}\deleted{compared} infer\deleted{ence}\deleted{s}
\added{\deleted{on}community structure} \added{when using data generated from:}
\deleted{based on}commercial catch, a\deleted{simulated} fixed-site sampling survey
design and the true \added{(simulated)} underlying populations.
\deleted{W}\deleted{\added{In doing so w}}\added{W}e \added{thereby} \deleted{i)}establish the potential
limitations of fishery-dependent data in providing a robust \replaced{characterisation}{picture} of
spatiotemporal distributions.\deleted{; and then ii) simulated an area closure
	based on areas defined from the different data sources at a range of
	temporal and spatial resolutions.}\deleted{Our framework allows users
	to explore the assumptions in modelling observational data and evaluate
	the underlying dynamics of such approaches at a fine spatial and
	temporal scale. In application to a mixed fishery exploiting four
	different populations w} \deleted{\added{W}e found d}\added{D}ifferent spatial patterns
were evident and \deleted{that }the effectiveness of the spatial closure reduced when
data were aggregated across larger spatial areas. \deleted{; and then
	ii)}\deleted{\added{We then simulated an area closure based on areas defined
	from the different data sources at a range of temporal and spatial
	resolutions.}} \added{A simulated area closure showed that} \replaced{a}{A}ggregation across time periods has less of a negative
impact on the closure success than \added{aggregation} over space. \deleted{ and w}\added{W}hile not
as effective as when based on on the true population, closures based on high
catch rates observed in commercial data were still able to reduce fishing on a
protected species. \\

\added{\replaced{MixFishSim}{Our framework} allows users to explore the assumptions in modelling
	observational data and evaluate the underlying dynamics of such
	approaches at a fine spatial and temporal scale. }\added{From our
	application w}\deleted{W}e conclude that commercial data, while
containing bias, provide a useful tool for managing catches in mixed fisheries
if applied at the correct spatiotemporal scale. \\
%%
\end{abstract}

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time, using a Gaussian spatial
process to define suitable habitat for different populations. As the precise
locations of the fish are known directly rather than inferred from sampling
or commercial catch, we can use the population model to validate how inference
from fisheries-dependent and fisheries independent sampling relates to the real
community structure in a way we could not with real data. We ii) compare, at
different spatial and temporal aggregations, the real (simulated) population
distributions to samples from fisheries-dependent and fisheries independent \textcolor{blue}{[CM: hyphenate or don't both, i.e., fishery-independent]}
catches to test if these are a true reflection of the relative density of the
populations. We then iii) simulate a fishery closure to protect a species based
on different spatial and temporal data aggregations. \\

@@

A Discrete-event simulation (DES) model of a hypothetical fishery was developed
as a software package (\textit{MixFishSim}). The modular approach enabled
efficient computation by allowing for sub-modules implemented on time-scales
appropriate to capture the characteristic of the different processes (Figure
1). The following sub-modules were included to capture the full system: 1)
Population dynamics, 2) Recruitment dynamics, 3) Population movement, 4)
fishery dynamics.\\

Population dynamics \added{for any number of species, as chosen by the user}
operate on a daily time-step \added{(with recruitment occurring only during
	defined seasons for each population)}, while population movement occurs
on a weekly time-step, with the fishing module operating on a tow-by-tow basis
(i.e.  multiple events a day). \\

@@

The basic population level processes were simulated using a modified two-stage
Deriso-Schnute delay difference model which models the fish populations in
terms of aggregate biomass of recruits and mature components rather than
keeping track of individuals\citep{Deriso1980, Schnute1985, Dichmont2003}. A
daily time-step was chosen to discretise continuous population processes on a
biologically relevant and computationally tractable timescale. Population
biomass growth was modelled as a function of previous recruited biomass,
intrinsic population growth and recruitment functionally linked to the adult
population size.  Biomass for each cell $c$ was incremented each day $d$ as
follows (the full parameter list is detailed in Table \ref{tab: pop_variable}): 
\begin{equation}
	\begin{split}
	B_{c,d+1} = &\\
	& \left(1 + \rho \right) B_{c,d} \cdot \mathrm{e}^{-Z_{c,d}} - \rho \cdot
	\mathrm{e}^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& \left(B_{c,d-1} \cdot \mathrm{e}^{-Z_{c,d-1}} + Wt_{R-1} \cdot \left( \alpha_{d-1} \cdot
	R_{\tilde{y}\left(c\right)}\right) \right)
	\hspace{0.9cm} + \\
	& Wt_{R} \cdot \left( \alpha_{d} \cdot R_{\tilde{y} \left(c\right)}
	\right) \\
	\end{split}
\end{equation}
where $\rho$ is Brody's coefficient, shown to be equal to $\mathrm{e}^{-K}$
when $K$ is the growth rate from a von Bertalanffy logistic growth model
\citep{Schnute1985}. $Wt_{R-1}$ is the average weight of fish prior to
recruitment, while $Wt_{R}$ is the average recruited weight. $\alpha_{d}$
represents the proportion of fish recruited during that day for the year, while
$R_{c,\tilde{y}(c)}$ is the annual recruits in year $y$ for cell $c$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} \cdot \left(1 -
	\mathrm{e}^{-(F_{c,d} + M_{c,d})}\right) \cdot B_{c,d}
\end{equation}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch is the sum of
those across all fleets and vessels, $C_{c, d} =
\sum\limits_{fl=1}^{FL}\sum\limits_{v=1}^{V} E_{fl, v, c, d} \cdot Q_{fl} \cdot
D_{c, d}$ with $fl$ and $FL$ the fleet and total number of fleets, $v$ and $V$
the vessel and total number of vessels respectively and $E_{fl,v,c,d}$ and
$Q_{fl}$ fishing effort and catchability of the gear, and $D_{c,d}$ is the
density of the population at the location fished. \\

@@

Recruitment is modelled as a function of adult biomass. In
\textit{MixFishSim}, it can either take the form of a stochastic Beverton-Holt
stock recruitment relationship, or a stochastic Ricker stock recruitment
relationship. The Beverton-Holt relationship is defined as\citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha \cdot S_{c,d})}{(\beta + S_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes. The stochastic Ricker form \citep{Ricker1954} is:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} \cdot \mathrm{e}^{(\alpha - \beta \cdot B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases.

@@

Population movement is a combination of directed (advective) movement where at
certain times of year the population moves towards spawning grounds by
increasing the probabilities of moving into the spawning grounds from adjacent
cells, and random (diffusive) movement, governed by a stochastic process where
movement between adjacent cells is described by a set of probabilities
\added{which are affected by the suitability of habitat, temperature in a cell
	and the thermal tolerance of a population to that temperature}.\\

\added{The combined process results in a population structure and movement
pattern unique to each population, with population movement occurring on a weekly
basis. The decision to model population movement on a weekly timescale was to
reflect that fish tend to aggregate in species specific locations that have
been observed to last around one to two weeks \citep{Poos2007}. Therefore this
process approximated the demographic shifts in fish populations throughout a
year with seasonal spawning patterns }(Figure \ref{S-fig:5}).

To simulate fish population distribution in space and time a Gaussian spatial
process was employed to model habitat suitability for each of the populations
on a 2d grid.  We first defined a Gaussian random field process, $\{S(c) : c
\in \mathbb{R}^2\}$, where for any set of cells $c_{1}, \dots, c_{n}$, the
joint distribution of $S = \{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian
with a \textit{Matérn} covariance structure, where the correlation strength
weakens with distance controlled by two parameters, with $\nu$ a scale
parameter in the units of distance and $\kappa$ a shape parameter which
determines the smoothness of the process. \added{We use the most commonly used
	Matérn covariance structure as it is a flexible form that under certain
	conditions is of the same form as an exponential function and
	it}\deleted{This} enables us to model the spatial autocorrelation
observed in animal populations where density is more similar in nearby
locations, \added{but that correlation decreases non-linearly}
\citep{Tobler1970, F.Dormann2007, Poos2007}. We change the parameters to
implement different spatial structures for the different populations using the
\textit{RandomFields} R package \citep{Schlater2015}. We define a stationary
habitat field with an anisotropic pattern (to simulate a depth gradient) and
combine it with a temporally dynamic thermal tolerance field to imitate two key
drivers of population dynamics without modelling the processes explicitly. Each
population was initialised at a single location, and subsequently moved across
the entire space according to a probabilistic distribution based on habitat
suitability (represented by the normalised values from the GRFs), temperature
tolerance and distance from current cell: 
\begin{equation}
	Pr(C_{wk+1} = J | C_{wk} =I) = \frac{\mathrm{e}^{-\lambda \cdot d_{I,J}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda \cdot d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{I,J}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{c,p}$ is the index of habitat
suitability for cell $c$ and population $p$, with $Tol_{c,p,wk}$ the
temperature tolerance for cell $c$ by population $p$ in week $wk$ (see below).
\\

@@

A time-varying temperature covariate changes the interaction between time and
suitable habitat on a weekly time-step. Each population $p$ was assigned a
thermal tolerance with mean, $\mu_{p}$ and standard deviation, $\sigma_{p}$ so that
each cell and population temperature tolerance is defined that:
\begin{equation}
	Tol_{c, p, wk} = \frac{1}{ \sqrt (2\pi \cdot \sigma^2_{p})} \cdot \exp(-
		\frac{(T_{c, wk} - \mu_{p})^2}{2 \cdot \sigma^2_{p}} )	
\end{equation}
Where $Tol_{c, p, wk}$ is the tolerance of population $p$ for cell $c$ in week
$wk$, $T_{c, wk}$ is the temperature in the cell given the week and $\mu_{p}$
and $\sigma_{p}$ the mean and standard deviation of the population temperature
tolerance. The variables, their meaning and units for population movement is
provided in Table \ref{tab:pop_move}. \\

@@

Each fleet of \textit{n} vessels was characterised by both a general
efficiency, $Q_{fl}$, and a population specific efficiency, ${Q_{fl, p}}$ which
are each bound by [0,1]. The product of these parameters [$Q_{fl} \cdot
Q_{fl, p}$] affects the overall catch rates for the fleet and the preferential
targeting of one species over another. This, in combination with the
parameter choice for the step-function defined below (as well as some
randomness from the exploratory fishing process) determined the preference of
fishing locations for the fleet.

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by relating recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$);
\begin{equation}
Rev_{c,d} = \sum_{p=1}^{P} L_{c,d,p} \cdot Pr_{p} 
\end{equation}
where $L_{c,d,p}$ is landings of a population $p$, and $Pr_{p}$ price of a
population. All population prices were kept the same across fleets and seasons.
Here, when fishing is successful vessels remain in a similar location and
continue to exploit the local fishing grounds. When unsuccessful, they move
some distance away from the current fishing location.  The movement distance
retains some degree of stochasticity, that can be controlled separately, but is
determined by the relationship: 
\begin{equation}
	Le = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		\left(log\left(\frac{\beta_{1}}{\beta_{3}}\right)\right) \cdot Rev
} 
\end{equation}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
($x_{t}$,$y_{t}$) to ($x_{t+1}$, $y_{t+1}$) is defined by:
\begin{equation}
	\begin{split}
 (x_{t+1}, y_{t+1}) =  & x_{t} + Le \cdot \cos \left(\frac{\pi \cdot
		 Br_{t+1}}{180}\right), \\ & y_{t} + Le \cdot \sin
 \left(\frac{\pi \cdot Br_{t+1}}{180}\right) \\	
 when  \hspace{0.5cm}     & Br_{t} < 180, Br_{t+1} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t} > 180, Br_{t+1} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $Le$ is the step length, $Br_t$ is the bearing at time $t$, $k$ the
concentration parameter from the von Mises distribution that we correlate with
the revenue so that $k = (Rev + 1 / RefRev) \cdot max_{k}$, where $max_{k}$ is
the maximum concentration value, $k$, and $RefRev$ is parametrised as for
$\beta_{3}$ in the step length function. Details of the variables, meaning and
units for fleet dynamics are provided in Table \ref{tab:fleet_dyn}.

@@

\subsection{Fisheries independent survey}

@@

We calibrated the simulation model for four example populations with different
demographics, growth rates, natural mortality and recruitment (Table
\ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and temperature
(Figures \ref{S-fig:3}, with temperature tolerance \ref{S-fig:4}) defined to be
unique to each population resulting in differently weekly distribution patterns
(Figures \ref{S-fig:5}-\ref{S-fig:7}). In addition, each of the populations was
assumed to have two defined spawning areas that result in the populations
moving towards these areas in pre-defined weeks (Figure \ref{S-fig:2}) with
population-specific movement rates (Table \ref{tab:1}).  \added{The population
demographics were defined to broadly represent three mobile low-medium value
groundfish species and one high value species with low mobility, with the
dynamics hypothetical but as you might expect to find in a typical demersal
fishery.}

@@

The fleets were calibrated to reflect five different characteristic fisheries
with unique exploitation dynamics (Table \ref{tab:2}). By setting different
catchability coefficients ($Q_{fl, p}$) we create different targeting
preferences between the fleets and hence spatial dynamics.  The learned random
walk process implies that within a fleet different vessels have different
spatial distributions based on individual experience. The step function was
calibrated dynamically within the simulations as the maximum revenue
obtainable was not known beforehand. This was implemented so that vessels take
smaller steps when fishing at a location that yields landings value in the top
90th percentile of the value experienced in that year so far (as defined per
fleet in Table \ref{tab:2}). \\

Fishing locations were chosen based on random search and, with increasing
proportion as time progressed, experience of profitable catches built up in the
same month from previous years and from the previous trip.  `Profitable' in
this context was defined as the locations where the top 70 \% of expected
profit would be found given revenue from previous trips and cost of movement to
the new fishing location. This probability was based on a logistic sigmoid
function with a lower asymptote of 0 and upper asymptote of 0.95, and a growth
rate that ensures the upper asymptote (where decisions are mainly based on past
knowledge) is reached approximately halfway through the simulation.  \\

@@

\section*{Appendices}

>>>>

\usepackage[final]{changes}

@@

%%
To understand how data resolution impacts inference on mixed fisheries
interactions we developed a highly resolved spatiotemporal discrete-event
simulation model \emph{MixFishSim} incorporating: i) delay-difference
population dynamics, ii) population movement using Gaussian Random Fields to
simulate patchy, heterogeneously distributed \added{and moving fish}
populations, and iii) fishery dynamics for multiple fleet characteristics based
on population targeting under an explore-exploit strategy.\deleted{ This is
	implemented via a mix of correlated random walk movement (for
	exploration) and learned behaviour (for exploitation) phases of the
	fisheries. }\\ 
%%
\deleted{Fifty years of sub-daily fishing activity was simulated and used to
	draw inference on the underlying community structures. W}\added{We
	applied \emph{MixFishSim} to\deleted{draw}}\deleted{compared}
infer\deleted{ence}\deleted{s} \added{\deleted{on}community structure}
\added{when using data generated from:} \deleted{based on}commercial catch,
a\deleted{simulated} fixed-site sampling survey design and the true
\added{(simulated)} underlying populations.  \deleted{W}\deleted{\added{In
		doing so w}}\added{W}e \added{thereby} \deleted{i)}establish
the potential limitations of fishery-dependent data in providing a robust
\replaced{characterisation}{picture} of spatiotemporal distributions.\deleted{;
	and then ii) simulated an area closure based on areas defined from the
	different data sources at a range of temporal and spatial
	resolutions.}\deleted{Our framework allows users to explore the
	assumptions in modelling observational data and evaluate the underlying
	dynamics of such approaches at a fine spatial and temporal scale. In
	application to a mixed fishery exploiting four different populations w}
\deleted{\added{W}e found d}\added{D}ifferent spatial patterns were evident and
\deleted{that }the effectiveness of the spatial closure reduced when data were
aggregated across larger spatial areas. \deleted{; and then
	ii)}\deleted{\added{We then simulated an area closure based on areas
		defined from the different data sources at a range of temporal
		and spatial resolutions.}} \added{A simulated area
	closure showed that} \replaced{a}{A}ggregation across time periods has
less of a negative impact on the closure success than \added{aggregation} over
space. \deleted{ and w}\added{W}hile not as effective as when based on on the
true population, closures based on high catch rates observed in commercial data
were still able to reduce fishing on a protected species. \\

\added{Our framework allows users to explore the assumptions in modelling
	observational data and evaluate the underlying dynamics of such
	approaches at a fine spatial and temporal resolutions. From our
	application w}\deleted{W}e conclude that commercial data, while
containing bias, provide a useful tool for managing catches in mixed fisheries
if applied at the correct spatiotemporal scale. \\
%%
\end{abstract}

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time, using a Gaussian spatial
process to define suitable habitat for different populations. As the precise
locations of the fish are known directly rather than inferred from sampling or
commercial catch, we can use the population model to validate how inference
from fisheries-dependent and fisheries-independent sampling relates to the real
community structure in a way we could not with real data. We ii) compare, at
different spatial and temporal aggregations, the real (simulated) population
distributions to samples from fisheries-dependent and fisheries-independent
catches to test if these are a true reflection of the relative density of the
populations. We then iii) simulate a fishery closure to protect a species based
on different spatial and temporal data aggregations. \\

@@

A Discrete-event simulation (DES) model of a hypothetical fishery was developed
as a software package (\textit{MixFishSim}). The modular approach enabled
efficient computation by allowing for sub-modules implemented on time-scales
appropriate to capture \deleted{the }characteristic\added{s} of the different processes (Figure
1). \replaced{S}{The following s}ub-modules \deleted{were included }to capture the full system \added{comprised}: 1)
\replaced{p}{P}opulation dynamics, 2) \replaced{r}{R}ecruitment dynamics, 3) \replaced{p}{P}opulation movement, 4)
fishery dynamics.\\

Population dynamics \added{for any number of species, as chosen by the user,}
operate on a daily time-step \added{(with recruitment occurring only during
	defined seasons for each population)}, while population movement occurs
on a weekly time-step, with the fishing module operating on a tow-by-tow basis
(i.e.\added{,} multiple events a day). \\

@@

The basic population level processes were simulated using a modified two-stage
Deriso-Schnute delay difference model \replaced{that}{which} models the fish populations in
terms of aggregate biomass of recruits and mature components rather than
keeping track of individuals \citep{Deriso1980, Schnute1985, Dichmont2003}. A
daily time-step was chosen to discretise continuous population processes on a
biologically relevant and computationally tractable timescale. Population
biomass growth was modelled as a function of previous recruited biomass,
intrinsic population growth and recruitment functionally linked to the adult
population size.  Biomass for each cell $c$ was incremented each day $d$ as
follows (\replaced{see}{the full parameter list is detailed in } Table \ref{tab: pop_variable} \added{for all parameter details}): 
\begin{equation}
	\begin{split}
	B_{c,d+1} = &\\
	& \left(1 + \rho \right) B_{c,d} \cdot \mathrm{e}^{-Z_{c,d}} - \rho \cdot
	\mathrm{e}^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& \left(B_{c,d-1} \cdot \mathrm{e}^{-Z_{c,d-1}} + W_{R-1} \cdot \left( \alpha_{d-1} \cdot
	R_{\tilde{y}\left(c\right)}\right) \right)
	\hspace{0.9cm} + \\
	& W_{R} \cdot \left( \alpha_{d} \cdot R_{\tilde{y} \left(c\right)}
	\right) \\
	\end{split}
\end{equation}
where $\rho$ is Brody's coefficient\textcolor{blue}{(Check)}, shown to be equal
to $\mathrm{e}^{-K}$ when $K$ is the \deleted{growth }rate \added{at which the
	asymptote is approached} from a von Bertalanffy \deleted{logistic
}growth model \citep{Schnute1985}. $W_{R-1}$ is the average weight of fish
prior to recruitment, while $W_{R}$ is the average recruited weight.
$\alpha_{d}$ represents the proportion of fish recruited during that day for
the year, while $R_{c,\tilde{y}(c)}$ is the annual recruits in year $y$ for
cell $c$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} \cdot \left(1 -
	\mathrm{e}^{-(F_{c,d} + M_{c,d})}\right) \cdot B_{c,d}
\end{equation}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch is the sum of
those across all fleets and vessels, $C_{c, d} =
\sum\limits_{fl=1}^{FL}\sum\limits_{v=1}^{V_{fl}} E_{fl, v, c, d} \cdot Q_{fl}
\cdot D_{c, d}$ with $fl$ and $FL$ the fleet and total number of fleets, $v$
and $V_{fl}$ the vessel and total number of vessels per fleet respectively and
$E_{fl,v,c,d}$ and $Q_{fl}$ fishing effort and catchability of the gear, and
$D_{c,d}$ is the density of the population at the location fished. \\

@@

Recruitment is modelled as a function of adult biomass. In
\textit{MixFishSim}, it can either take the form of a stochastic Beverton-Holt
stock recruitment relationship, or a stochastic Ricker stock recruitment
relationship. The Beverton-Holt relationship is defined as\citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha \deleted{\cdot} S_{c,d})}{(\beta + S_{c,d})} \\
	     \added{\ln(}R_{c,d}\added{)} \sim & \deleted{\log} N[(\replaced{\ln}{\log}(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes. The stochastic Ricker form \citep{Ricker1954} is:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} \cdot \mathrm{e}^{(\alpha - \beta \cdot B_{c,d})} \\	
	\ln(R_{c,d}) \sim & N[(\ln(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the
density-dependent reduction in productivity as the SSB increases.

@@

Population movement is a combination of directed (advective) movement where at
certain times of year the population moves towards spawning grounds by
increasing the probabilities of moving into the spawning grounds from adjacent
cells, and random (diffusive) movement, governed by a stochastic process where
movement between adjacent cells is described by a set of probabilities.
\added{Stochastic probabilities are affected by the suitability of habitat,
	temperature in a cell and the thermal tolerance of a population to that
	temperature}.\\

\added{The combined process results in a population structure and movement
	pattern unique to each population, with population movement occurring
	on a weekly basis. \replaced{M}{The decision to m}odel\added{ing}
	population movement on a weekly timescale\deleted{ was to}
	reflect\added{s} that fish tend to aggregate in
	species\added{-}specific locations\deleted{ that have been} observed to
	last \replaced{between}{around } one \replaced{and}{to} two weeks
	\citep{Poos2007}. Therefore this process approximated the demographic
	shifts in fish populations throughout a year with seasonal spawning
	patterns }(Figure \ref{S-fig:5}).

To simulate fish population distribution in space and time a Gaussian spatial
process was employed to model habitat suitability for each of the populations
on a 2d grid.  We first defined a Gaussian random field process, $\{S(c) : c
\in \mathbb{R}^2\}$, where for any set of cells $c_{1}, \dots, c_{n}$, the
joint distribution of $S = \{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian
with a \textit{Matérn} covariance structure, where the correlation strength
weakens with distance controlled by two parameters, with $\nu$ a scale
parameter in the units of distance and $\kappa$ a shape parameter which
determines the smoothness of the process. \added{We use the most commonly used
	Matérn covariance structure as it is a flexible form that \replaced{contains the}{under certain
	conditions is of the same form as an} exponential \added{and double exponential as special cases}\deleted{ function} and
	it}\deleted{This} enables us to model the spatial autocorrelation
observed in animal populations where density is more similar in nearby
locations\deleted{, \added{but that correlation decreases non-linearly}}
\citep{Tobler1970, F.Dormann2007, Poos2007}. We change the parameters to
implement different spatial structures for the different populations using the
\textit{RandomFields} R package \citep{Schlater2015}. We define a stationary
habitat field with an anisotropic pattern (to simulate a depth gradient) and
combine it with a temporally dynamic thermal tolerance field to imitate two key
drivers of population dynamics without modelling the processes explicitly. Each
population was initialised at a single location, and subsequently moved across
the entire space according to a probabilistic distribution based on habitat
suitability (represented by the normalised values from the GRFs), temperature
tolerance and distance from current cell: 
\begin{equation}
	Pr(C_{wk+1} = J | C_{wk} =I) = \frac{\mathrm{e}^{-\lambda \cdot d_{I,J}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda \cdot d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{I,J}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{c,p}$ is the index of habitat
suitability for cell $c$ and population $p$, with $Tol_{c,p,wk}$ the
temperature tolerance for cell $c$ by population $p$ in week $wk$ (see below).
\\

@@

A time-varying temperature covariate changes the \deleted{interaction between time and}
suitable habitat on a weekly time-step. Each population $p$ was assigned a
thermal tolerance with mean, $\mu_{p}$ and standard deviation, $\sigma_{p}$ so that
each cell and population temperature tolerance is defined \replaced{as}{that}:
\begin{equation}
	Tol_{c, p, wk} = \frac{1}{ \sqrt{2\pi  \sigma^2_{p}}}  \exp \left(-
		\frac{(T_{c, wk} - \mu_{p})^2}{2  \sigma^2_{p}} \right)	
\end{equation}
Where $Tol_{c, p, wk}$ is the tolerance of population $p$ for cell $c$ in week
$wk$, $T_{c, wk}$ is the temperature in the cell given the week and $\mu_{p}$
and $\sigma_{p}$ the mean and standard deviation of the population temperature
tolerance \replaced{(see Table \ref{tab:pop_move} for variable descriptions).}{. The variables, their meaning and units for population movement is
provided in Table \ref{tab:pop_move}}. \\

@@

Each fleet of \textit{$n_{fl}$} vessels was characterised by both a general
efficiency, $Q_{fl}$, and a population specific efficiency, ${Q_{fl, p}}$ which
are each bound by [0,1]. The product of these parameters [$Q_{fl} \cdot Q_{fl,
	p}$] affects the overall catch rates for the fleet and the preferential
targeting of one species over another. This, in combination with the parameter
choice for the step-function defined below (as well as some randomness from the
exploratory fishing process) determined the preference of fishing locations for
the fleet.

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by relating recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$);
\begin{equation}
Rev_{c,d} = \sum_{p=1}^{P} L_{c,d,p} \cdot Pr_{p} 
\end{equation}
where $L_{c,d,p}$ is landings of a population $p$, and $Pr_{p}$ price of a
population. All population prices were kept the same across fleets and seasons.
Here, when fishing is successful vessels remain in a similar location and
continue to exploit the local fishing grounds. When unsuccessful, they move
some distance away from the current fishing location.  The movement distance
retains some degree of stochasticity, that can be controlled separately, but is
determined by the relationship: (\textcolor{blue}{[CM: use $\ln$ where we use the natural logarithm. Double-check this equation is right as per the code. As it looks different here: \url{https://github.com/pdolder/MixFishSim/blob/master/R/step_length.R]}. There it looks to simplify to $b_2(b_1/b_2)^{Rev/b_3}$, assuming parameters are positive.})
\begin{equation}
	Le = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		\left(log\left(\frac{\beta_{1}}{\beta_{3}}\right)\right) \cdot Rev
} 
\end{equation}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
($x_{t}$,$y_{t}$) to ($x_{t+1}$, $y_{t+1}$) is defined by:
\begin{equation}
	\begin{split}
 (x_{t+1}, y_{t+1}) =  & x_{t} + Le \cdot \cos \left(\frac{\pi \cdot
		 Br_{t+1}}{180}\right), \\ & y_{t} + Le \cdot \sin
 \left(\frac{\pi \cdot Br_{t+1}}{180}\right) \\	
 when  \hspace{0.5cm}     & Br_{t} < 180, Br_{t+1} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t} > 180, Br_{t+1} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $Le$ is the step length, $Br_t$ is the bearing at time $t$, $k$ the
concentration parameter from the von Mises distribution that we correlate with
the revenue so that $k = (Rev + 1 / RefRev) \cdot max_{k}$, where $max_{k}$ is
the maximum concentration value, $k$, and $RefRev$ is parametrised as for
$\beta_{3}$ in the step length function. Details of the variables, meaning and
units for fleet dynamics are provided in Table \ref{tab:fleet_dyn}.

@@

\subsection{Fisheries-independent survey}

@@

We calibrated the simulation model for four example populations with different
demographics, growth rates, natural mortality and recruitment (Table
\ref{tab:1}). Habitat preference (Figure \ref{S-fig:1}) and temperature
(Figures \ref{S-fig:3}, with temperature tolerance \ref{S-fig:4}) defined to be
unique to each population resulting in differently weekly distribution patterns
(Figures \ref{S-fig:5}-\ref{S-fig:7}). In addition, each of the populations was
assumed to have two defined spawning areas that result in the populations
moving towards these areas in pre-defined weeks (Figure \ref{S-fig:2}) with
population-specific movement rates (Table \ref{tab:1}).  \added{The population
demographics were chosen to broadly represent three mobile low-medium value
groundfish species and one high value species with low mobility, with the
dynamics hypothetical but \deleted{as you} might \added{be} expect\added{ed} \deleted{to find }in a typical demersal
fishery.}

@@

\replaced{F}{The f}leets were calibrated to reflect five different characteristic fisheries
with unique exploitation dynamics (Table \ref{tab:2}). By setting different
catchability coefficients ($Q_{fl, p}$) we create different targeting
preferences between the fleets and hence \added{different} spatial dynamics.  The learned random
walk process implies that within a fleet different vessels have different
spatial distributions based on individual experience. The step function was
calibrated dynamically within the simulations as the maximum revenue
obtainable was not known beforehand. This was implemented so that vessels take
smaller steps when fishing at a location that yields landings value in the top
90th percentile of the value experienced in that year so far (as defined per
fleet in Table \ref{tab:2}). \\

Fishing locations were chosen based on random search and, with increasing
proportion as time progressed, experience of profitable catches built up in the
same month from previous years and from the previous trip.  `Profitable' in
this context was defined as the locations where the top 70 \% of expected
profit would be found given revenue from previous trips and cost of movement to
the new fishing location. This probability was based on a logistic sigmoid
function with a lower asymptote of 0 and upper asymptote of 0.95, and a \replaced{slope}{growth
rate} that ensures the upper asymptote (where decisions are mainly based on past
knowledge) is reached approximately halfway through the simulation.  \\

@@

\newpage
